Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-14 surfaces-write-up/surface-write-up.tex
added some examples for ghost points

\DeclareMathOperator{\supp}{Supp}
\DeclareMathOperator{\initial}{in_\prec}
\DeclareMathOperator{\gin}{gin}
\DeclareMathOperator{\Eff}{Eff}
\DeclareMathOperator{\sat}{sat}
\DeclareMathOperator{\newspan}{span}
\DeclareMathOperator{\proj}{Proj}
\DeclareMathOperator{\spec}{Spec}
\DeclareMathOperator{\Te}{T_=}
\DeclareMathOperator{\Tp}{T_+}
\DeclareMathOperator{\Tm}{T_-}
\DeclareMathOperator{\num}{p}
\DeclareMathOperator{\den}{q}
%\captionsetup[table]{belowskip = 4pt}

@@

This example easily generalizes to the following situation: Suppose $D = \sum_{i=1}^{n} \frac{d}{s}$
\end{example}

\begin{example}
\label{eg:radical}
In this example, we justify the necessity of adding ghost points for the bounds obtained in Lemma ~\ref{P-m-generators}, which are also necessary when the divisors are not all contained in a hyperplanes.

 As an example for why we need to add in ghost points for $\bp^2$, consider $\frac{-1}{5}V(x_0^2) + \frac{1}{7}V(x_1^2) + \frac{1}{17}V(x_2^2)$. In degree $595 = 5* 7 * 17$, this has dimension $6$, but we cannot possibly use the lattice approach, as all lower degrees only have dimension 1. In fact, we may construct similar examples using any $x_i^m$ for $f_i$ with appropriate coefficients.
\end{example}
\todo{construct class of additional examples}
\todo{The example above needs to be changed because $x_0^2$ can be replaced by $2x_0$}

>>>>

\DeclareMathOperator{\supp}{Supp}
\DeclareMathOperator{\initial}{in_\prec}
\DeclareMathOperator{\gin}{gin}
\DeclareMathOperator{\Eff}{Eff}
\DeclareMathOperator{\sat}{sat}
\DeclareMathOperator{\newspan}{span}
\DeclareMathOperator{\proj}{Proj}
\DeclareMathOperator{\spec}{Spec}
\DeclareMathOperator{\Te}{T_=}
\DeclareMathOperator{\Tp}{T_+}
\DeclareMathOperator{\Tm}{T_-}
\DeclareMathOperator{\num}{p}
\DeclareMathOperator{\den}{q}
\DeclareMathOperator{\lcm}{lcm}
%\captionsetup[table]{belowskip = 4pt}

@@

This example generalizes slightly to any divisor of the form $D = \frac{1}{k}H_0 - \frac{1}{k+1}H_1 \in \di \bp^2 \otimes \bq,$ with $k \in \bn$, showing that there will always exist a generator in degree $a(a+1).$

This example further generalizes to the following situation: Suppose $D = \sum_{i=1}^{n} \frac{\num_i}{\den_i}D_i \in \di \bp^m \otimes \bq,$ where $\deg D_i = \bida_i$ and $\deg D = \frac{1}{\lcm_{1 \leq i \leq n}(\den_i \cdot \bida_i)}$. Then, if $D_i = V(f_i)$ where all $f_i$ can be written as a polynomial function in $x_0, \ldots, x_{m-1},$ we have $R_D$ always has a generator in degree $\lcm(\den_i \cdot \bida_i)$.
\end{example}

\begin{rem}
\label{rem:ghost-motivation}
Example ~\ref{eg:hyperplane} shows that the naive generalization of ~\ref[Theorem 8]{dorney:canonical} of generation in degree at most $\sum_{i=1}^{n}\ell_i$ cannot possibly hold. The reason for this is that the divisors may ``be contained in a hyperplane,'' i.e., be expressable as functions in $m$ of the $m+1$ variables on $\bp^m$. This problem can easily be circumvented by adding in ``ghost points.'' That is, we may add divisors of the form $0 \cdot H_i$ to $D$, and reorder so that if $D = \sum_{i=1}^{n}\alpha_i V(f_i)$, then $f_0, \ldots, f_n$
\end{rem}

\begin{example}
\label{eg:radical}
In this example, we show that it is still, in general, necessary to add ghost points, even when the irreducible components of a divisor are not all contained in a hyperplane.

Consider $\frac{-1}{5}V(x_0^2 + x_1^2 + x_2^2) + \frac{1}{7}V(x_0^2 + x_1^2 + x_3^2) + \frac{1}{17}V(x_0^2 + x_2^2 + x_3^2) + \frac{1}{596}V(x_1^2 + x_2^2 + x_3^2)$. In degree $355216 = 5* 7 * 17 * 596$, this has dimension $10$. 
\todo{Clarify what this example means by the lattice approach}
However, we possibly use the lattice approach, as all lower degrees only have dimension 1. In fact, we may construct similar examples using any set of degree $n$ divisor with very small degree.
\end{example}
\todo{construct class of additional examples}
\todo{The example above needs to be changed because $x_0^2$ can be replaced by $2x_0$}
